Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions

### 2015-03-12 HGmodelMANUSCRIPT/HGmodel_draft3.tex
Fixed typo in equation (1)
Added definitions of M, F, Sigma and color scheme to FF ranking Fig.

\subsection{Order parameters from experiments}\label{expORDp}
\todo{DONE
%We should mention the sign issue and refer to the blog post:
%http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html somewhere.
%Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments 
%and simulations" or similar, and also move the "forking" explanation here.
}
The order parameter of a hydrocarbon C--H vector is defined as 
\begin{equation}\label{orderP}
S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle,
\end{equation} 
where the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
splitting with $^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
The absolute values of order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
values between 20.2--22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
For a more detailed discussion see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
\todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.
COMMENT: There is also the option to put the most relevant blog posts as a supplementary material. The advantage of this would
be that the posts would be peer reviewed as well.}.

@@

To get a general idea of the quality of the glycerol backbone and choline headgroup structures in different models, we calculated 
the order parameters for these parts from thirteen different lipid models (Table~\ref{systems}) and 
plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
Two criteria were used to judge the quality of the model: there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, the {\bf magnitude}
should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
The results for each force field in respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
\todo{DONE}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonTable.pdf}
  \caption{\label{FullHydrationComparisonTable}
  Ranking of force fields based on data of Fig.~\ref{HGorderparameters}. 
  Markus, please describe the definition of M, F and $\Sigma$ here.
  } 
\end{figure}

>>>>

\subsection{Order parameters from experiments}\label{expORDp}
\todo{DONE
%We should mention the sign issue and refer to the blog post:
%http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html somewhere.
%Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments 
%and simulations" or similar, and also move the "forking" explanation here.
}
The order parameter of a hydrocarbon C--H vector is defined as 
\begin{equation}\label{orderP}
S_{{\rm CH}}=\frac{1}{2}\langle 3 \cos^2 \theta-1 \rangle,
\end{equation} 
where the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
splitting with $^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
The absolute values of order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
values between 20.2--22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
For a more detailed discussion see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
\todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.
COMMENT: There is also the option to put the most relevant blog posts as a supplementary material. The advantage of this would
be that the posts would be peer reviewed as well.}.

@@

To get a general idea of the quality of the glycerol backbone and choline headgroup structures in different models, we calculated 
the order parameters for these parts from thirteen different lipid models (Table~\ref{systems}) and 
plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
Two criteria were used to judge the quality of the model: there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, the {\bf magnitude}
should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
The results for each force field in respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
\todo{DONE}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonTable.pdf}
  \caption{\label{FullHydrationComparisonTable}
  Rough ranking of force fields based on data of Fig.~\ref{HGorderparameters}.
  "M" indicates a magnitude problem,
  "F" a forking problem.
  Letter size shows the level (0-4) of severity;
  the $\Sigma$-column shows the sum of these, i.e., the "total severity".
  Color scheme:
  "within experimental error" (dark green),
  "almost within experimental error" (light green),
  "clear deviation from experiments" (light red), and
  "major deviation from experiments" (dark red).
  } 
\end{figure}
